The effects of context inference on motor learning: savings, de-adaptation and spontaneous recovery

### 2021-09-23 adaptation/article/article.tex
Update on Overleaf.
Minor changes to abstract

\abstract{Abstract}
 Humans are able to adapt their movements to changes in the dynamics of the environment and of their own bodies, a phenomenon known as motor adaptation. Motor adaptation has been shown to be contextual in experiments in which participants must learn to adapt to different modified dynamics, and must be able to switch between the learned adaptations dynamically. In real life, the information that we use to determine how to perform a task is often incomplete or misleading, which affects our ability to switch between tasks. In this work, we focus on the behavioral phenomena effected by imperfect contextual information and present a hierarchical computational model for motor adaptation based on exact Bayesian inference, which explicitly takes into account contextual information and how the dynamics of context inference affect learning and action selection. We show how our model provides a unifying explanation for the effects of sensory cues and proprioceptive information on switching between tasks, the effects of environmental volatility and predictability on learning, and the effects of previously-learned adaptations on switching between tasks, in addition to the well studied savings, de-adaptation and spontaneous recovery.

@@

It has been shown that humans can learn multiple tasks during the course of a single experiment, as well as learn to adapt to often-opposing modifications to the dynamics of movements \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference can occur between tasks \citep[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}, which slows down learning. Additionally, humans have been shown to dynamically switch between different learned tasks \citep[e.g.][]{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}. The mechanisms behind these switches, the specifics of how these switches manifest in behavior, and the ways in which switching can be impeded or improved, are still an active topic of research.

>>>>

\abstract{Abstract}
Humans have been shown to adapt their movements when a sudden change to the dynamics of the environment is introduced, a phenomenon called motor adaptation. If the change is reverted, the adaptation is also quickly reverted. Human are also able to adapt to multiple changes in dynamics presented separately, and to be able to switch between adapted movements on the fly. Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations. In this work, we focus on the behavioral phenomena effected by imperfect contextual information and present a hierarchical computational model for motor adaptation based on exact Bayesian inference, which explicitly takes into account contextual information and how the dynamics of context inference affect learning and action selection. We show how our model provides a unifying explanation for the effects of sensory cues and proprioceptive information on switching between tasks, the effects of environmental volatility and predictability on learning, and the effects of previously-learned adaptations on switching between tasks, in addition to the well studied savings, de-adaptation and spontaneous recovery.

@@

It has been shown that humans can learn multiple tasks during the course of a single experiment, as well as learn to adapt to often-opposing modifications to the dynamics of movements \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference between tasks can slow down learning \citep[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}. Additionally, humans have been shown to dynamically switch between different learned adaptation \citep[e.g.][]{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}. The mechanisms behind these switches, the specifics of how these switches manifest in behavior, and the ways in which switching can be impeded or improved, are still an active topic of research.
